---
title: "Model Fitting (JAGS)"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Model Fitting (JAGS)}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r}
#| include = FALSE
jcutils::knitr_default_setup(
  echo = TRUE, collapse = TRUE, comment = "#>", fig.asp = 0.65, base_size = 12
)
library(MixMod)
library(mcmcr)
library(rjags)
```


## Backend

### Motivation

Fitting models in JAGS (specifically, `rjags`) requires three distinct steps.

1. model compilation, `rjags::jags.model()`;

1. warm-up sampling, `update()`;

1. extracting posterior samples, `rjags::jags.samples()`.

This package makes `complete_jags_fit` available, a wrapper around these steps
that also allows a progress bar in `knitr`.

As such, control of a JAGS sampler is done using a list of specifically named 
controls that can be validated with `validate_jags_n()`.

```{r validate-jags-n}
#| results = "hide"
jags_n <- 
  list(
    n.adapt  = 1000,  # Adaptation
    n.update = 1000,  # Warm-up
    n.iter   = 10000, # Monitoring
    n.thin   = 1,     # Thinning interval
    n.chains = 4      # Number of chains
  )
```


### Example

The following example shows how such a fit can be made. Note the final object
is a list with these elements.

- `data`. List originally passed to JAGS;

- `model`. Object returned from `jags.model`;

- `output`. Posterior samples stored in an `mcmcr-object`;

- `jags_n`. List originally passed to JAGS.

```{r jags-fit}
jags_file <- 
  textConnection(
    "
    model {
      for (i in 1:n) { x[i] ~ dnorm(mu,tau) }
      mu ~ dnorm(0, 0.1)
      tau ~ dgamma(3, 3)
    }
    "
  )

jags_data <-
  list(x = rnorm(100, mean = 2, sd = 1/sqrt(4)), n = 100)

jags_n <- 
  list(n.adapt = 100, n.update = 100, n.iter = 2e3, n.thin = 1, n.chains = 1)

jags_fit <- 
  MixMod::complete_jags_fit(
    data = jags_data,
    file = jags_file,
    jags_n = jags_n,
    varnames = c("mu", "tau")
)
```

We can obtain posterior estimates by simply printing the `mcmcr` object. For
more information on dealing with the output, see the **MCMC Functionality**
vignette.

```{r jags-fit-output-show}
print(jags_fit$output)
```

