---
title: "MCMC Functionality"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{MCMC Functionality}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r}
#| include = FALSE
jcutils::knitr_default_setup(
  echo = TRUE, collapse = TRUE, comment = "#>", fig.asp = 0.65, base_size = 12
)
library(MixMod)
library(mcmcr)
library(rjags)
library(magrittr)
```


## Working with MCMC samples.


### Compact (`mcmcr`)

One of the main libraries we leverage to manipulate `rjags` output is `mcmcr`.
All samples are held in a `mcmcr` object that serves as a list of arrays.

```{r mcmcr-show}
mcmcr::mcmcr_example
```

Note that the default summaries are the posterior median and chains are 
collapsed in the calculation.

```{r mcmcr-beta-estimate}
mcmcr::estimates(mcmcr::mcmcr_example$beta, fun = median)
```

We extend this functionality to allow for estimates on a per-chain basis.

```{r mcmcr-beta-estimate-per-chain}
MixMod::estimates_by_chain(mcmcr::mcmcr_example$beta, fun = mean)
```

Further, we also implement `purrr`-like mapping functions that do not have the 
restriction that input and output dimensions should match.
As with the inspiration, there are three versions:
`map_mcmcarray()`, `map2_mcmcarray()` and `pmap_mcmcarray()`.

For example, for some 2-length vector `alpha` and 2-by-2 matrix `beta` we can
return another `mcmcr-object`, that is in this case a scalar parameter.

```{r mcmcr-beta-map}
MixMod::map2_mcmcarray(
  mcmcr::mcmcr_example$alpha,
  mcmcr::mcmcr_example$beta,
  function(a, b) as.numeric(a %*% b %*% a)
)
```


### Tidy (`draws`)

While `mcmcr` is compact, manipulating draws using `tidyverse` and `ggplot2`
functionality is not straightforward. 
[Tidybayes](http://mjskay.github.io/tidybayes/) aims to "make it easy to 
integrate such systems into a tidy data and ggplot workflow".

We build from these principles and typically convert MCMC objects into a
**draws** format, see '*Working with Draws*' section for an example.

The following methods provide utility for creating **draws** data frames.

- `gather_draws_by_varnames()`.
  Obtain draws in `tidy` format using `varname` convention.

- `is_valid_mcmcr_args()`.
  Checks for valid `mcmcr` object and parameter (`varnames`) combination.

- `create_tidybayes_specs()`.
  Creates expression(s) for use with `tidybayes` package.

- `as.mcmc.list.mcmcr()`.
  Required for `tidybayes` extension.

One can easily join draws with truth in simulation studies using this simple S3
generic.

- `to_long_format()`.
  Converts an R object to a data frame with one row per value.
    
```{r to-long-format}
x <- matrix(1:6, nrow = 3, ncol = 2)
to_long_format(x)
```


## Working with Draws

A **draws** like data frame has one row for each MCMC draw and columns 
`.chain`, `.iteration`, `.draw`, `.term` and `.value`.

```{r draws-show}
draws <- 
  MixMod::gather_draws_by_varnames(
    x = mcmcr::mcmcr_example,
    varnames = "beta"
  )

print(draws)
```

All are self-explanatory but note that `.term` is a special data type where 
the variable name and dimensions are verified and accessible using the `term`
package.

Manipulating these draws can be done via:

- `draws_split_by_chain()`.
  Splits one data frame into a list of draws, one for each chain.

- `draws_summarise()`.
  Calculate mean and credible intervals of draws using `ggdist::point_interval`.

- `draws_rename_terms()`.
  Rename the `.term` column without affecting dimension, useful for plotting.
  
- ~~`format_matrix_draws_summary()`.
  Creates data frame (to be passed to `kbl`) that mirrors matrix shape.~~

- `draws_unnest_term()`.
  Add new columns representing index values of each dimension.


## Visualising Draws

Visualisation is typically done by nesting a **draws** data frame, creating 
what we call a **draws_vis** data frame.

See the following for an example involving a grouped **draws** object and
anonymous plotting function.

```{r draws-vis}
draws_vis <- 
  MixMod::draws_vis_by_group(
    draws = dplyr::group_by(draws, .data$.term),
    .gg = ~ MixMod::ggdraws_trace(.x)
  )

print(draws_vis)
```


### Individual Plots (`ggdraws`)

Various plotting methods that convert **draws** to a `ggplot2` plot are 
available as follows.
    
- `ggdraws_dens()`.
  Density plot.
    
- `ggdraws_trace()`.
  Trace plot.
    
- `ggdraws_acf()`.
  Autocorrelation plots from results arising from `stats::acf()`.
  
- `ggdraws_group_trace()`.
  Tile plot using colour to represent cluster membership.
  
- `ggdraws_prediction()`.
  Time-series plot of given `.truth` alongside estimates.


### Multiple Plots (`patchwork`)

We can organise the plots within a **draws-vis** table as a patchwork. This
is typically done by grouping where two groups represents the rows and columns
respectively

Methods acting on the `plot` columns directly are listed as follows.

- `draws_vis_align_limits()`.
  Align x-axis or y-axis across all plots in table.

- `draws_vis_add_subtitle()`.
  Add subtitle using concatenated output of other columns


## Worked Example

Consider the following example that brings this all together.

```{r mcmcr-to-trace}
# Start with mcmcr object
mcmcr::mcmcr_example %>%
  
  # Only obtain the draws for beta[1,1], ..., beta[2,2]
  gather_draws_by_varnames("beta") %>%
  
  # Discard draws not from the first chain
  dplyr::filter(.data$.chain == 1) %>%
  
  # Create plot on a term-by-term basis
  MixMod::draws_vis_by_group(~ MixMod::ggdraws_dens(.x)) %>%
  
  # Align y-axis and x-axis for better comparisons
  MixMod::draws_vis_align_limits() %>%
  
  # Group by two dimensions of beta (with arbitrary labels)
  MixMod::draws_unnest_term(i = c("var1", "var2"), j = c("VAR1", "VAR2")) %>%
  dplyr::group_by(i, j) %>%
  
  # Rename the term and add subtitles
  MixMod::draws_rename_terms(pattern = "beta", replacement = "Beta") %>%
  MixMod::draws_vis_add_subtitle(.term, i, j) %>%
  
  # Create grid from grouping structure
  MixMod::patchwork_grid()
```
