---
title: "RJAGS Example (Gamma Random Sample)"
format: html
toc: true
---

```{r setup}
#| include: false
library(mcmcrutils)
```

## JAGS File

We obtain the filepath to an example JAGS model specification and store
    it in `jags_file` to later use.
    
The content of this file is as follows.

```{r setup-file}
#| include: false
jags_file <- 
  system.file(
    "model", "gamma-random-sample.jags",
    package = "mcmcrutils",
    mustWork = TRUE
  )
```

```{embed}
#| file: !expr jags_file
#| filename: !expr sprintf("JAGS file (%s)", basename(jags_file))
#| lang: r
```

## JAGS Controls

To fit the model, the sampling specification must be set.

```{r setup-jags-n}
jags_n <-
  list(
    n.adapt = 1e4,
    n.update = 1e4,
    n.iter = 5e4,
    n.thin = 5,
    n.chains = 4
  )
```

## Model Fitting

Then, the model can be fit in one function call with a progress bar displayed 
    for each stage.
    
```{r complete-jags-fit}
#| message: false
fit <-
  mcmcrutils::complete_jags_fit(
    file = jags_file,
    data = list(x = rgamma(200, 2, 3)),
    jags_n = jags_n,
    varnames = c("alpha", "lambda", "y")
  )
```

## Analysis

**TODO**. Requires development into `ggdraws` and `patchgrid`.
